Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
pf <- read.csv("pseudo_facebook.tsv",sep = '\t')
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()+
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender)))+
  geom_line(aes(color = gender),stat = 'summary', fun.y = median)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

age_gender_groups <- group_by(subset(pf, !is.na(gender)),age,gender)

pf.fc_by_age_gender <- summarise(age_gender_groups,
                          mean_friend_count = mean(friend_count),
                          median_friend_count = median(friend_count),
                          n = n())
ungroup(pf.fc_by_age_gender)
pf.fc_by_age_gender <- arrange(pf.fc_by_age_gender, age, gender)

head(pf.fc_by_age_gender)
tail(pf.fc_by_age_gender)

pf.fc_by_age_gender<- pf%>%
  filter(!is.na(gender))%>%
  group_by(age, gender)%>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n())%>%
  ungroup()%>%
  arrange(age)

ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender)+
  geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)

```


***

### Ratio Plot
Notes:

```{r Ratio Plot}

ggplot(aes(x = age, y = female/male),
       data = pf.fc_by_age_gender.wide)+
  geom_line()+
  geom_hline(yintercept = 1, linetype = 2)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014-pf$tenure/(365))
head(pf)
summary(pf$year_joined)
table(pf$year_joined)
?cut

pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks =c(2004,2009,2011,2012,2014),
                             include.lowest = FALSE, right = TRUE)

head(pf)
table(pf$year_joined.bucket)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color = year_joined.bucket),stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color = year_joined.bucket),stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary',fun.y = mean, linetype =2)


```

***

### Friending Rate
Notes:

```{r Friending Rate}

users <- subset(pf, pf$tenure>0)

users$rate <- users$friend_count/users$tenure

summary(users$rate)

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

ggplot(aes(y = friendships_initiated/tenure, x = tenure),
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
  

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)



ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

yo <- read.csv('yogurt.csv')
str(yo)

yo$id <- factor(yo$id)

ggplot(aes(price), data = yo)+ geom_histogram(binwidth = 10)

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

yo$all.purchases = (yo$strawberry+yo$blueberry+yo$pina.colada+yo$plain+yo$mixed.berry)

```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x= time, y = price), data = yo)+
  geom_jitter(alpha = 1/4, shape =21, fill = I('#F79420'))

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids<-sample(levels(yo$id),16)

ggplot(aes(x=time, y=price),data =subset(yo, id %in% sample.ids))+
  facet_wrap(~id)+
  geom_line()+
  geom_point(aes(size = all.purchases), pch =1)

set.seed(3690)
sample.ids<-sample(levels(yo$id),16)

ggplot(aes(x=time, y=price),data =subset(yo, id %in% sample.ids))+
  facet_wrap(~id)+
  geom_line()+
  geom_point(aes(size = all.purchases), pch =1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

```{r scatterplots}
install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[,c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])

```
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)

library(reshape2)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

