Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv',sep = '\t')

qplot(x = age, y = friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x=age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) +
  xlim(13,90) +
  scale_y_sqrt()
  
ggplot(aes(x=age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
  

summary(pf$age)
```

***
#the data looks bimodular. Looks like Skewed with long tail
### Overplotting
Notes:

```{r Overplotting}

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

plot = ggplot(aes(x = age, y = friendships_initiated), data = pf)
plot + geom_point(alpha = 1/10, position = 'jitter') +
  xlim(13,90)
plot + geom_point(alpha = 1/20) +
  xlim(13,90)+
  coord_trans(y = "sqrt")
```

***

### Overplotting and Domain Knowledge
Notes: 

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

age_groups <- group_by(pf,age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(pf.fc_by_age)

pf.fc_by_age <- pf%.%
  group_by(age)%.%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %.%
  arrange(age)

plot = ggplot(aes(x = age, y =friend_count_mean), data = pf.fc_by_age)+
  geom_line()

plot+geom_line()+xlim(13,90)

```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
plot = ggplot(aes(x = age, y = friend_count), data = pf)

plot2 = plot + geom_point(alpha = 1/20, position = position_jitter(h=0),
                  color = 'orange') +
  xlim(13,90)+
  coord_trans(y = "sqrt")+
  geom_line(stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary', fun.y = quantile,  fun.args=list(probs=0.1),
             linetype = 2, color = 'blue')+
  geom_line(stat = 'summary', fun.y = quantile,  fun.args=list(probs=0.5),
             color = 'blue')+
  geom_line(stat = 'summary', fun.y = quantile,  fun.args=list(probs=0.9),
            linetype =2, color = 'blue')


plot2+coord_cartesian(xlim = c(13,40), ylim = c(0,1000))

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

cor(pf$age, pf$friend_count, method = "pearson")

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with( subset(pf, pf$age <=70), cor.test(age, friend_count, method = "pearson"))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

plot = ggplot(aes(y = likes_received, x = www_likes_received), data= pf) +geom_point()

plot
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

plot+xlim(0, quantile(pf$www_likes_received,0.95))+
  ylim(0,quantile(pf$likes_received,0.95))+
  geom_smooth(method = 'lm', color = 'red')

cor.test(pf$www_likes_received, pf$likes_received, method = "pearson")


```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}



```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)

```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x= Month, y = Temp), data= Mitchell)+geom_point()+
  scale_x_discrete(breaks = seq(0,203,12))

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

coord_cartesian()
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

ggplot(aes(x= age, y = friend_count_mean), data = pf.fc_by_age)+ geom_line()

```

***

### Age with Months Means

```{r Age with Months Means}

pf$age_with_months <- (pf$age + (pf$dob_month/12))
  
head(pf$dob_month/12, 20)

  
age_month_groups <- group_by(pf,age_with_months)
pf.fc_by_age_months <- summarise(age_month_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
```

Programming Assignment
```{r Programming Assignment} 

p1 <- ggplot(aes(x = age, y =friend_count_mean), 
           data = subset(pf.fc_by_age, age<71))+
  geom_line()+
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months<71))+
  geom_line()+
  geom_smooth()

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count), 
       data = subset(pf, age<71))+
  geom_line(stat = 'summary',fun.y = mean)

library(gridExtra)
grid.arrange(p2,p1,p3,ncol=1)

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}


```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

